---
title: R Script to reroduce results of ;Global, regional and country epidemiology
  and prevalence of stunting, wasting and underweight in low- and middle- income countries,
  2006-2018
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

# clear R global environment
rm(list=ls())

#Load the libraries

```{r, message=FALSE}
# load libraries
library(olsrr) # linear regression
library(tidyverse) # general stuff
library(hrbrthemes) # pretty plots
library(ggplot2) # pretty plots
library(plotly) # interactive plots
library(gapminder) # interactive plots
library(pastecs) # stat.desc
library(metafor) # for meta-analysis: mixed-effects logistic and Poisson regression models
library(meta) # meta-regression, GLMM, forest plots,(subgroup) meta-analyses.
library(sp) # spatial data
library(rgdal) #  projection/transformation operations for shapefiles
library(sf) # Simple Features for R
library(rnaturalearth)
library(tmap) # cool maps
library(ggpubr) # 'ggplot2' Based Publication Ready Plots
library(ggpmisc) #Miscellaneous Extensions to 'ggplot2'
library(spData)
library(cowplot)# plot grid
```

# load data

world
```{r}
dat <- read.csv(file="DHS77_META.csv", header=T,stringsAsFactors=F)
hdi<- read.csv(file="HDI.csv", header=T,stringsAsFactors=F)
educ<-read.csv(file="Women_Education.csv")
dat <- merge(dat,hdi, by=c("Country"), all.x=T)
dat <- merge(dat,educ, by=c("Country"), all.x=T)

#dat_meta <- merge(dat,hdi, by=c("Country"), all.x=T)
dat$name=dat$Country

```



# display maps

```{r map stunt}
WorldData <- map_data('world')
WorldData %>% filter(region != "Antarctica") -> WorldData
WorldData <- fortify(WorldData)

p <- ggplot()
p <- p + geom_map(data=WorldData, map=WorldData,
                  aes(x=long, y=lat, group=group, map_id=region),
                  fill="white", colour="#7f7f7f", size=0.5)
p <- p + geom_map(data=dat, map=WorldData,
                  aes(fill=stunt, map_id=Country),
                  colour="#7f7f7f", size=0.5)
p <- p + coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90))
#p <- p + scale_fill_continuous(low="thistle2", high="darkred", guide="colorbar")
p <-p+ scale_fill_distiller(palette = "Spectral")

p <- p + scale_y_continuous(breaks=c())
p <- p + scale_x_continuous(breaks=c())
p <- p + labs(fill="Stunting(%)", title="", x="", y="")
p <- p + theme_bw()
p <- p + theme(panel.border = element_blank())

par(mai=c(0,0,0.2,0),xaxs="i",yaxs="i")
p 

```




```{r}
world_stunt = world %>%
  filter(continent!= "Antarctica", !is.na(iso_a2)) %>%
  #filter(continent == "Africa", !is.na(iso_a2)) #%>%
  left_join(worldbank_df, by = "iso_a2") %>%
  dplyr::select(name, subregion, gdpPercap, HDI, pop_growth) %>%
  left_join(dat, by = "name") %>%
  st_transform("+proj=longlat +datum=WGS84 +no_defs")

worldbank_df
# Map the output
tm_shape(world_stunt) + tm_polygons(col="HDI_2018", style="quantile") +
  tm_legend(outside=TRUE)
```


# stunting meta-analysis: Figure 1

```{r stunting}
# Fit model
m1.pft <- metaprop(stunt_n, Height_Age_N, data = dat, studlab = paste (Country, Year, sep = ", "), comb.fixed =F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")
m1.plogit <- update(m1.pft, sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m1.plogits<- update(m1.pft, byvar= dat$UN.Sub.Region,print.byvar=F)

# plot forest plot
pdf("figs/Figure2.pdf", width = 12, height = 25)
forest(m1.plogits,
       leftlabs = c("Country, Year of survey", "Stunted", "Total"),
       digits = 2,
       col.by="black",
       #sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of stunting (%)",
       xlim = c(0, 60), at = c(0,20,40,60),pscale = 100,
       backtransf = T)
dev.off()

```


# wasting meta-analysis: Figure 2

```{r wasting}
# Fit model

m2.pft <- metaprop(wasted_n, Weight_Height_N, data = dat, studlab = paste (Country, Year, sep = ", "),comb.fixed=F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")
m2.plogit <- update(m2.pft , sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m2.plogits<- update(m2.pft, byvar= dat$UN.Sub.Region,print.byvar=F)

# plot forest plot
pdf("figs/Figure3.pdf", width = 12, height = 25)
forest(m2.plogits,
       leftlabs = c("Country, Year of survey", "Wasted", "Total"),
       digits = 2,
       col.by="black",
       sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of wasting (%)",
       xlim = c(0, 25), at = c(0,5,10,15,20,25),pscale = 100,
       backtransf = T)
dev.off()

```


# underweight meta-analysis: Figure 3

```{r underweight}
# Fit model

m3.pft <- metaprop(underweight_n, Weight_Age_N, data = dat, studlab = paste (Country, Year, sep = ", ") ,comb.fixed=F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")

m3.plogit <- update(m1.plogit , sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m3.plogits<- update(m3.pft, byvar= dat$UN.Sub.Region,print.byvar=F)

# plot forest plot
pdf("figs/Figure4.pdf", width = 12, height = 25)
forest(m3.plogits,
       leftlabs = c("Country, Year of survey", "Underweight", "Total"),
       digits = 2,
       col.by="black",
       sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of underweight (%)",
       xlim = c(0, 50), at = c(0,10,20,30,40,50),pscale = 100,
       backtransf = T)
dev.off()

```


# stunting meta-analysis: Supplementary Figure S2

```{r stunting S2}
# Fit model
m1.pft <- metaprop(stunt_n, Height_Age_N, data = dat, studlab = paste (Country, Year, sep = ", "), comb.fixed =F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")
m1.plogit <- update(m1.pft, sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m1.plogits<- update(m1.pft, byvar= dat$UN.Regions,print.byvar=F)

# plot forest plot
pdf("figs/SupplementaryFigureS5.pdf", width = 12, height = 18)
forest(m1.plogits,
       leftlabs = c("Country, Year of survey", "Stunted", "Total"),
       digits = 2,
       col.by="black",
       #sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of stunting (%)",
       xlim = c(0, 60), at = c(0,20,40,60),pscale = 100,
       backtransf = T)
dev.off()

```

# wasting meta-analysis: Supplementary Figure S3

```{r wasting S3}
# Fit model

m2.pft <- metaprop(wasted_n, Weight_Height_N, data = dat, studlab = paste (Country, Year, sep = ", "),comb.fixed=F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")
m2.plogit <- update(m2.pft , sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m2.plogits<- update(m2.plogit, byvar= UN.Regions,print.byvar=F)

# plot forest plot
pdf("figs/SupplementaryFigureS6.pdf", width = 12, height = 18)
forest(m2.plogits,
       leftlabs = c("Country, Year of survey", "Wasted", "Total"),
       digits = 2,
       col.by="black",
       sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of wasting (%)",
       xlim = c(0, 25), at = c(0,5,10,15,20,25),pscale = 100,
       backtransf = T)
dev.off()

```



# Underweight meta-analysis: Supplementary Figure S4

```{r underweight S4}
# Fit model

m3.pft <- metaprop(underweight_n, Weight_Age_N, data = dat, studlab = paste (Country, Year, sep = ", ") ,comb.fixed=F,
                   sm = "PFT", pscale = 100, method.ci = "NAsm",
                   method.tau = "DL")

m3.plogit <- update(m3.pft, sm = "PLOGIT",backtransf = T)
# subgroup analysis by UN subregion
m3.plogits<- update(m3.plogit, byvar= UN.Regions,print.byvar=F)

# plot forest plot
pdf("figs/SupplementaryFigureS7.pdf", width = 15, height = 22)
forest(m3.plogits,
       leftlabs = c("Country, Year of survey", "Underweight", "Total"),
       digits = 2,
       col.by="black",
       sortvar = Year,
       squaresize=0.5,
       col.square="navy",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       print.pval.Q = T,
       xlab="Prevalence of underweight (%)",
       xlim = c(0, 50), at = c(0,10,20,30,40,50),pscale = 100,
       backtransf = T)
dev.off()

```



#Meta-regression. Fitting generalized linear mixed effects models for stunting

```{r stunting meta-regression}
# Use glmm to estimate the risk of stunting
model_S <- rma.glmm(xi=stunt_n, ni=Height_Age_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+dat$HDI_Cat2, dat=dat, method="ML")
round(exp(coef(summary(model_S))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_S,digits=3)

# add malaria but subset to SSA
model_S <- rma.glmm(xi=stunt_n, ni=Height_Age_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+HDI_Cat2+scale(Mal_Prev), dat=dat,subset = SSA == "Sub-Saharan Africa", method="ML")
round(exp(coef(summary(model_S))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_S,digits=3)

```


Meta-regression. Fitting generalized linear mixed effects models for wasting

```{r wasting meta-regression}
# Use glmm to estimate the risk of wasting
model_W <- rma.glmm( xi=wasted_n, ni=Weight_Height_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+dat$HDI_Cat2, dat=dat, method="ML")
round(exp(coef(summary(model_W))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_W,digits=3)

# add malaria but subset to SSA
model_W <- rma.glmm(xi=wasted_n, ni=Weight_Height_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+HDI_Cat2+scale(Mal_Prev), dat=dat,subset = SSA == "Sub-Saharan Africa", method="ML")
round(exp(coef(summary(model_W))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_W,digits=3)

```

Meta-regression. Fitting generalized linear mixed effects models for underweight

```{r underweight meta-regression}

# Use glmm to estimate the risk of underweight
model_U <- rma.glmm( xi=underweight_n, ni=Weight_Age_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+dat$HDI_Cat2, dat=dat, method="ML")
round(exp(coef(summary(model_U))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_U,digits=3)

# add malaria but subset to SSA
model_U <- rma.glmm(xi=underweight_n, ni=Weight_Age_N, measure="PLO", mods = ~ relevel(factor(UN.Sub.Region),ref="Central Asia")+HDI_Cat2+scale(Mal_Prev), dat=dat,subset = SSA == "Sub-Saharan Africa", method="ML")
round(exp(coef(summary(model_U))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)
print(model_U,digits=3)

```







# plot for HDI: linear trend + confidence interval

```{r}
# plot stunting as function of HDI

s1=ggplot(dat, aes(x=HDI_2018, y=stunt)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

s1=s1+ xlab("Human development index")+
  ylab('Prevalence of stunting (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
s1
cor.test(dat$HDI_2018, dat$stunt, method=c("spearman"))

# plot wasting

w1=ggplot(dat, aes(x=HDI_2018, y=wasted)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

w1=w1+ xlab("Human development index")+
  ylab('Prevalence of wasting (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
w1
cor.test(dat$HDI_2018, dat$wasted, method=c("spearman"))

# plot underweight

u1=ggplot(dat, aes(x=HDI_2018, y=underweight)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

u1=u1+ xlab("Human development index")+
  ylab('Prevalence of underweight (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
u1

cor.test(dat$HDI_2018, dat$underweight, method=c("spearman"))

```

# plot for Malaria: linear trend + confidence interval

```{r}
# plot stunting as function of malaria

s1=ggplot(subset(dat, UN.Regions %in% c("Africa")), aes(x=Mal_Prev, y=stunt)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()


s1=s1+ xlab("Malaria prevalence (%)")+
  ylab('Prevalence of stunting (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
s1
cor.test(dat$Mal2, dat$stunt, method=c("spearman"))

# plot wasting

w1=ggplot(subset(dat, UN.Regions %in% c("Africa")), aes(x=Mal_Prev, y=wasted)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

w1=w1+ xlab("Malaria prevalence (%)")+
  ylab('Prevalence of wasting (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
w1

cor.test(dat$Mal2, dat$wasted, method=c("spearman"))
# plot underweight

u1=ggplot(subset(dat, UN.Regions %in% c("Africa")), aes(x=Mal_Prev, y=underweight)) +
  geom_point(alpha=0.8, size=3) +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

u1=u1+ xlab("Malaria prevalence (%)")+
  ylab('Prevalence of underweight (%)')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 20),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=20),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
u1

cor.test(dat$Mal2, dat$underweight, method=c("spearman"))

```

```{r}
# Use row names as point labels
#df$name <- rownames(df)
ed=ggscatter(dat, x = "HDI_2018", y = "Value", palette = "jco", add = "reg.line", conf.int = TRUE,
   label = "Country", repel = TRUE)+
  stat_cor(aes(), method = "spearman", label.y = 90)+
  theme_ipsum()

ed=ed+ xlab("Human Development Index")+
  ylab('Education Attainment for Women')+
  theme(
    axis.title.x=element_text(angle=0, color='black',face = "bold",size = 14),
    axis.title.y=element_text(angle=90, color='black', face='bold', size=14),
    axis.text.x = element_text(face="bold", color="black",size=20, angle=0),
    axis.text.y = element_text(face="bold", color="black", size=20, angle=0),
    legend.text = element_text(colour="black", size=20,face="bold"),
    legend.title = element_text(colour="black", size=20,face="bold"))
ed

```

